---
title: "WUENIC Talking Points: `r params$country`"
output:
  pdf_document:
    latex_engine: xelatex
geometry: "left=1cm,right=1cm,top=1.1cm,bottom=1.1cm"
fontsize: 12pt
params:
  country: "Default"
header-includes:
  - \usepackage{enumitem}
  - \setlist[itemize,1]{label=\textbullet}
  - \setlist[itemize,2]{label=\(\circ\)}
  - \setlist[itemize,3]{label=\(\infty\)}
  - \setmainfont{Arial}
  - \usepackage{hyperref}
  - \hypersetup{colorlinks=true, linkcolor=blue, urlcolor=blue, breaklinks=true}
---

```{r setup, include=FALSE}
options(scipen = 999)

# filter to selected country and its region peers
df <- wuenic_dta %>% filter(country == params$country, lvl_1 == "country")
hpv_df <- hpv_dta %>% filter(country == params$country, lvl_1 == "country")

# pull region and iso3c for this country
ctry_iso   <- unique(df$iso3c)
ctry_regn  <- unique(df$Region)
ctry_name  <- unique(df$country)

# regional peers (all countries in same region, for comparisons)
reg_peers <- wuenic_dta %>% filter(Region == ctry_regn, lvl_1 == "country")
hpv_peers <- hpv_dta %>% filter(Region == ctry_regn, lvl_1 == "country")

# subnational data (admin1 within country) -- update lvl_2 value to match your data
sub_df <- wuenic_dta %>% filter(iso3c == ctry_iso)

wuenic_src <- str_glue("Source: WHO/UNICEF Estimates of National Immunization Coverage, {rev_yr} revision")
```

```{r setup-vax, include=FALSE}
# key vaccines and years
vax <- df %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1", "mcv2"), year %in% c(comp_yr, rev_yr - 1, rev_yr))

# unvaccinated labels
unvac_lbl <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated_lbl) %>%
  pivot_wider(names_from = year, values_from = unvaccinated_lbl) %>%
  rename_with(~paste0(., "_lbl"), .cols = (ncol(.) - 2):ncol(.))

# unvaccinated raw values
unvac <- vax %>%
  select(Region, iso3c, country, year, vaccine, unvaccinated) %>%
  pivot_wider(names_from = year, values_from = unvaccinated) %>%
  left_join(., unvac_lbl) %>%
  distinct() %>%
  mutate(pchng = round((.[[7]] / .[[5]] - 1) * 100, 0),
         pchng_txt = case_when(
           .[[8]] == .[[10]] ~ str_glue("remained relatively stable between {comp_yr} and {rev_yr} at {.[[10]]}"),
           pchng < 0 ~ str_glue("declined {abs(pchng)}% from {.[[8]]} in {comp_yr} to {.[[10]]} in {rev_yr}"),
           pchng > 0 ~ str_glue("increased {pchng}% from {.[[8]]} in {comp_yr} to {.[[10]]} in {rev_yr}"),
           TRUE ~ NA_character_))

# vaccinated
vacc <- vax %>%
  select(Region, iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = year, values_from = vaccinated) %>%
  mutate(pchng = round((.[[7]] / .[[5]] - 1) * 100, 1),
         pchng_txt = case_when(pchng > 0 ~ "increased", pchng < 0 ~ "decreased", TRUE ~ "remained the same")) %>%
  label_vals_millions(.[[5]], "vacc_comp_yr_lbl") %>%
  label_vals_millions(.[[7]], "vacc_rev_yr_lbl")

# coverage
cvg <- vax %>%
  select(Region, iso3c, country, year, vaccine, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage)

# target population (surviving infants), dtp1 only
tgt <- vax %>%
  filter(vaccine == "dtp1") %>%
  select(Region, iso3c, country, year, vaccine, target) %>%
  pivot_wider(names_from = year, values_from = target) %>%
  mutate(diff = .[[7]] - .[[5]],
         moreless = case_when(diff < 0 ~ "fewer", diff > 0 ~ "more", TRUE ~ "same"),
         to_vax = case_when(
           diff < 0 ~ "for vaccine coverage to increase, the number of children vaccinated needs to either increase or decline at a slower rate than the decline in surviving infant target population",
           diff > 0 & diff < 1000000 ~ "vaccinating approximately the same number of children as currently reached will be sufficient due to the anticipated decrease or relatively small increases in the birth cohort",
           diff > 0 & diff >= 1000000 ~ "maintaining current coverage requires vaccinating an increasing number of children, which will require substantial increases in immunization programme and health system capacity")) %>%
  label_vals(diff, "diff_lbl") %>%
  mutate(diff_abs = abs(diff)) %>%
  label_vals_millions(diff_abs, "diff_abs_lbl")
```

```{r setup-overview, include=FALSE}
# vaccines achieving >= 90% in this country in rev_yr
gt90 <- df %>%
  select(-target_grp) %>%
  distinct() %>%
  filter(year == rev_yr) %>%
  mutate(gt90 = ifelse(coverage >= 90, 1, 0))

n_vax_total <- n_distinct(gt90$vaccine)
n_vax_gt90  <- sum(gt90$gt90, na.rm = TRUE)

# coverage change prev_yr to rev_yr for the key vaccines
cvg_chng <- vax %>%
  filter(year %in% c(rev_yr, rev_yr - 1)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(diff = .[[5]] - .[[4]],
         diff_txt = case_when(
           diff > 0 ~ str_glue("increased from {.[[4]]}% in {rev_yr-1} to {.[[5]]}% in {rev_yr}"),
           diff < 0 ~ str_glue("declined from {.[[4]]}% in {rev_yr-1} to {.[[5]]}% in {rev_yr}"),
           TRUE ~ str_glue("was the same in {rev_yr} as it was in {rev_yr-1} ({.[[5]]}%)")))
```

```{r setup-progress-ia2030-zd-target, include=FALSE}
# zero-dose (unvaccinated dtp1) for this country (dtp1 is used as a proxy for the IA2030 goal)
zd <- df %>%
  filter(vaccine == "dtp1", year >= 2000) %>%
  select(iso3c, country, year, unvaccinated)

# ia2030 annual target (halve 2019 zero-dose by 2030)
zd_2019 <- zd %>% filter(year == 2019) %>% pull(unvaccinated)
ia2030_annual_target <- zd_2019 / 2

target <- zd %>%
  filter(year == 2019) %>% # baseline data only
  mutate(target_2030 = unvaccinated / 2, # defines the goal: 50% reduction in unvaccinated people by 2030
         factor = (unvaccinated - target_2030) / (2030 - 2019)) %>% # annual reduction rate needed: total drop divided by the 11-year gap
  crossing(year_new = 2020:2030) %>% # "join" that duplicates each country's row for every year from 2020 to 2030
  mutate(unvaccinated_new = unvaccinated - factor * (year_new - 2019)) %>% # applies linear formula to calculate the target number of zero-dose children
  select(iso3c, country, year = year_new, unvaccinated_new)

# text for ia2030 comparison in rev_yr
zd_rev_yr <- zd %>% filter(year == rev_yr) %>% pull(unvaccinated)
ia2030_rev_yr <- target %>% filter(year == rev_yr) %>% pull(unvaccinated_new)
zd_pcnt_diff  <- round((zd_rev_yr / ia2030_rev_yr - 1) * 100, 0)
zd_diff_txt   <- case_when(zd_pcnt_diff > 0 ~ "higher than", zd_pcnt_diff < 0 ~ "lower than", TRUE ~ "the same as")

zd_rev_yr_lbl     <- label_vals_millions_num(zd_rev_yr)
ia2030_rev_yr_lbl <- label_vals_millions_num(ia2030_rev_yr)
```

<!-- ===================== PAGE 1: OVERVIEW ===================== -->

\textcolor{cyan}{\Large Overview}

\vspace{1em}

- `r if (n_vax_gt90 == 0) "None" else n_vax_gt90` of the `r n_vax_total` childhood vaccines tracked for `r ctry_name` achieved coverage of 90% or higher in `r rev_yr`.

\vspace{1em}

- Coverage of DTP1 `r cvg_chng %>% filter(vaccine == "dtp1") %>% pull(diff_txt)`, DTP3 `r cvg_chng %>% filter(vaccine == "dtp3") %>% pull(diff_txt)`, and MCV1 `r cvg_chng %>% filter(vaccine == "mcv1") %>% pull(diff_txt)`.

\vspace{1em}

- In `r rev_yr`, the number of zero-dose children in `r ctry_name` (`r zd_rev_yr_lbl`) was approximately `r abs(zd_pcnt_diff)`% `r zd_diff_txt` the annual goal to achieve the IA2030 target of halving zero-dose children by 2030 (`r ia2030_rev_yr_lbl`).

\vspace{1em}

```{r setup-heatmap-overview, include=FALSE}

# filter hpv data to bind with other vaccines
hpv_merge <- hpv_df %>% 
  filter(vaccine_code == "PRHPVC_F") %>% # HPVc females only
  select(-vaccine) %>% 
  rename(vaccine = vaccine_code)

df <- df %>% 
  rbind(hpv_merge)

# heatmap for all vaccines, this country only
heatmap_ovw_df <- df %>%
  mutate(year = as.character(year),
         vaccine = case_when(
           vaccine == "hepbb" ~ "HepBB", vaccine == "hepb3" ~ "HepB3",
           vaccine == "hib3"  ~ "Hib3",  vaccine == "rotac" ~ "RotaC",
           vaccine == "menga" ~ "MengA", vaccine == "PRHPVC_F" ~ "HPVc",
           TRUE ~ toupper(vaccine))) %>%
  mutate(cvg_cat = case_when(
    coverage < 60 ~ "<60", coverage %in% 60:69 ~ "60-69",
    coverage %in% 70:79 ~ "70-79", coverage %in% 80:89 ~ "80-89",
    coverage %in% 90:94 ~ "90-94", coverage >= 95 ~ ">=95", TRUE ~ NA_character_))

heatmap_ovw_df$vaccine <- factor(heatmap_ovw_df$vaccine,
  levels = c("BCG", "HepBB", "DTP1", "DTP3", "Hib3", "HepB3", "PCVC", "RotaC", "POL3", "IPV1", "IPVC", "MCV1", "RCV1", "MCV2", "YFV", "MengA", "HPVc"))

heatmap_overview <-
  heatmap_ovw_df %>%
  ggplot(aes(x = year, y = vaccine, fill = coverage)) +
  geom_tile(color = "white") +
  #geom_text(data = heatmap_ovw_df %>% filter(coverage >= 60), aes(label = coverage), colour = "black", vjust = 0.5, size = 3) +
  geom_text(data = heatmap_ovw_df, aes(label = coverage), colour = "white", vjust = 0.5, size = 3) +
  labs(x = "", y = "",
       title = str_glue("Vaccine coverage (%), {ctry_name}, 2000–{rev_yr}"),
       caption = wuenic_src) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm")) +
  scale_fill_gradientn(
    colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
    values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
    guide = "colorbar", limits = c(0, 100),
    name = "Vaccine coverage (%)  ",
    breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-overview-heatmap, echo=FALSE, fig.width=10, fig.height=6.2}
print(heatmap_overview)
```

\newpage

<!-- ===================== PAGE 2: DTP1 ===================== -->

```{r setup-dtp1, include=FALSE}
v <- "dtp1"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

# dtp1 coverage for this country in rev_yr and comp_yr
ctry_cvg_rev  <- cvg_v %>% pull(7)   # rev_yr column
ctry_cvg_comp <- cvg_v %>% pull(5)   # comp_yr column

# regional rank (among peers) in rev_yr
reg_rank <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

n_peers <- reg_peers %>% filter(vaccine == v, year == rev_yr) %>% n_distinct("iso3c")

# # subnational variation
# sub_dtp1 <- sub_df %>%
#   filter(vaccine == v, year == rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
# 
# sub_variation_txt <- case_when(
#   sub_dtp1$range_cvg >= 20 ~ str_glue("Subnational variation is high, with coverage ranging from {sub_dtp1$min_cvg}% to {sub_dtp1$max_cvg}% across {sub_dtp1$n_admin1} admin1 areas"),
#   sub_dtp1$range_cvg >= 10 ~ str_glue("Subnational variation is moderate, with coverage ranging from {sub_dtp1$min_cvg}% to {sub_dtp1$max_cvg}% across {sub_dtp1$n_admin1} admin1 areas"),
#   TRUE ~ str_glue("Subnational variation is low, with coverage ranging from {sub_dtp1$min_cvg}% to {sub_dtp1$max_cvg}% across {sub_dtp1$n_admin1} admin1 areas"))

# dtp1->dtp3 dropout
dropout_dtp <- df %>%
  filter(year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1", "dtp3")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1 - dtp3) / dtp1) * 100, 0),
         dropout_txt = case_when(
           dropout < 5  ~ "low DTP dropout rates imply good ability to provide a complete vaccine series early in life",
           dropout <= 10 ~ "medium DTP dropout rates imply moderate ability to provide a complete vaccine series early in life",
           dropout > 10  ~ "high DTP dropout rates imply poor ability to provide a complete vaccine series early in life"))

dropout_comp_dtp <- dropout_dtp %>%
  select(-c(dtp1, dtp3, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(diff = .[[4]] - .[[3]],
         diff_txt = case_when(
           diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[4]]}%)"),
           diff < 0 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[4]]}%)"),
           TRUE ~ str_glue("the same as in {comp_yr}")))
```

\textcolor{cyan}{\Large DTP1: Reaching zero-dose children}

\vspace{1em}

- In `r rev_yr`, DTP1 coverage in `r ctry_name` was `r ctry_cvg_rev`%, compared to `r ctry_cvg_comp`% in `r comp_yr`.

\vspace{1em}

- In `r rev_yr`, the number of zero-dose children in `r ctry_name` (`r zd_rev_yr_lbl`) was approximately `r abs(zd_pcnt_diff)`% `r zd_diff_txt` the annual IA2030 target (`r ia2030_rev_yr_lbl`).

\vspace{1em}

- With `r tgt$diff_abs_lbl` `r tgt$moreless` surviving infants in `r rev_yr` compared to `r comp_yr`, `r tgt$to_vax`.

\vspace{1em}

- Among `r ctry_regn` region peers, `r ctry_name` ranked `r reg_rank$rank` out of `r n_peers` countries in DTP1 coverage in `r rev_yr` (`r ctry_cvg_rev`% vs. regional peers).

\vspace{1em}

\vspace{1em}

- DTP1-to-DTP3 dropout in `r rev_yr` was `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout)`%, `r dropout_comp_dtp %>% pull(diff_txt)`. `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout_txt) %>% str_to_sentence()`.

\vspace{1em}

```{r setup-heatmap-dtp1-sub, include=FALSE}
# # heatmap of dtp1 by admin1 over time
# heatmap_sub_dtp1 <-
#   sub_df %>%
#   filter(vaccine == "dtp1", year >= 2000) %>%
#   mutate(year = as.factor(year)) %>%
#   ggplot(aes(x = year, y = reorder(admin1_master, coverage), fill = coverage)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = coverage), colour = "black", size = 2.8) +
#   labs(x = "", y = "",
#        title = str_glue("DTP1 coverage (%), by admin1, {ctry_name}, 2000-{rev_yr}"),
#        caption = wuenic_src) +
#   scale_x_discrete(position = "top") +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
#         axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
#   scale_fill_gradientn(
#     colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
#     values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
#     guide = "colorbar", limits = c(0, 100),
#     name = "Vaccine coverage (%)  ",
#     breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-dtp1-sub, echo=FALSE, fig.width=10, fig.height=6.5}
#print(heatmap_sub_dtp1)
```

\newpage

<!-- ===================== PAGE 3: DTP3 ===================== -->

```{r setup-dtp3, include=FALSE}
v <- "dtp3"
unvac_v <- unvac %>% filter(vaccine == v)
cvg_v   <- cvg   %>% filter(vaccine == v)
vacc_v  <- vacc  %>% filter(vaccine == v)

ctry_cvg_rev_dtp3  <- cvg_v %>% pull(7)
ctry_cvg_comp_dtp3 <- cvg_v %>% pull(5)

# regional rank
reg_rank_dtp3 <- reg_peers %>%
  filter(vaccine == v, year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

# un- and undervaccinated (unvaccinated for dtp3)
unvac_dtp3_rev  <- unvac_v %>% pull(7)
unvac_dtp3_lbl  <- label_vals_millions_num(unvac_dtp3_rev)

# population growth context
tgt_chng <- tgt %>%
  mutate(pop_grwth = round((.[[7]] / .[[5]] - 1) * 100, 0)) %>%
  label_vals_millions(.[[7]], "tgt_rev_yr") %>%
  label_vals_millions(.[[5]], "tgt_comp_yr")

pop_chng_txt <- if (ctry_cvg_rev_dtp3 < ctry_cvg_comp_dtp3 & tgt[[7]] > tgt[[5]]) {
  str_glue("However, due to population growth of approximately {tgt_chng$pop_grwth}% between {comp_yr} and {rev_yr}, DTP3 coverage in {rev_yr} ({ctry_cvg_rev_dtp3}%) was lower than in {comp_yr} ({ctry_cvg_comp_dtp3}%)")
} else { "" }

# # subnational variation for dtp3
# sub_dtp3 <- sub_df %>%
#   filter(vaccine == v, year == rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
# 
# sub_var_dtp3_txt <- case_when(
#   sub_dtp3$range_cvg >= 20 ~ str_glue("Subnational variation is high, with coverage ranging from {sub_dtp3$min_cvg}% to {sub_dtp3$max_cvg}% across {sub_dtp3$n_admin1} admin1 areas"),
#   sub_dtp3$range_cvg >= 10 ~ str_glue("Subnational variation is moderate, with coverage ranging from {sub_dtp3$min_cvg}% to {sub_dtp3$max_cvg}% across {sub_dtp3$n_admin1} admin1 areas"),
#   TRUE ~ str_glue("Subnational variation is low, with coverage ranging from {sub_dtp3$min_cvg}% to {sub_dtp3$max_cvg}% across {sub_dtp3$n_admin1} admin1 areas"))
```

\textcolor{cyan}{\Large DTP3: A marker of routine immunization service delivery to children}

\vspace{1em}

- The number of children vaccinated with DTP3 `r vacc_v %>% pull(pchng_txt)` from `r vacc_v %>% pull(vacc_comp_yr_lbl)` in `r comp_yr` to `r vacc_v %>% pull(vacc_rev_yr_lbl)` in `r rev_yr`. `r pop_chng_txt`.

\vspace{1em}

- In `r rev_yr`, `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive DTP3. Dropout was `r dropout_comp_dtp %>% pull(diff_txt)`. `r dropout_dtp %>% filter(year == rev_yr) %>% pull(dropout_txt) %>% str_to_sentence()`.

\vspace{1em}

- In `r rev_yr`, DTP3 coverage in `r ctry_name` was `r ctry_cvg_rev_dtp3`%, leaving `r unvac_dtp3_lbl` children un- or undervaccinated.

\vspace{1em}

- Among `r ctry_regn` region peers, `r ctry_name` ranked `r reg_rank_dtp3$rank` out of `r n_peers` countries in DTP3 coverage in `r rev_yr`.

\vspace{1em}

\vspace{1em}

```{r setup-heatmap-dtp3-sub, include=FALSE}
# heatmap_sub_dtp3 <-
#   sub_df %>%
#   filter(vaccine == "dtp3", year >= 2000) %>%
#   mutate(year = as.factor(year)) %>%
#   ggplot(aes(x = year, y = reorder(admin1_master, coverage), fill = coverage)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = coverage), colour = "black", size = 2.8) +
#   labs(x = "", y = "",
#        title = str_glue("DTP3 coverage (%), by admin1, {ctry_name}, 2000-{rev_yr}"),
#        caption = wuenic_src) +
#   scale_x_discrete(position = "top") +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
#         axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
#   scale_fill_gradientn(
#     colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
#     values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
#     guide = "colorbar", limits = c(0, 100),
#     name = "Vaccine coverage (%)  ",
#     breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-dtp3-sub, echo=FALSE, fig.width=10, fig.height=6.5}
#print(heatmap_sub_dtp3)
```

\newpage

<!-- ===================== PAGE 4: MCV ===================== -->

```{r setup-mcv, include=FALSE}
v_mcv <- c("mcv1", "mcv2")
unvac_v <- unvac %>% filter(vaccine %in% v_mcv)
cvg_v   <- cvg   %>% filter(vaccine %in% v_mcv)

# mcv1 coverage change
cvg_mcv1 <- cvg_v %>%
  filter(vaccine == "mcv1") %>%
  mutate(diff = .[[7]] - .[[5]],
         cvg_chng_txt = case_when(
           diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1 percentage point of the {comp_yr} level"),
           diff < -1 ~ str_glue("was lower than in {comp_yr} ({.[[5]]}%)"),
           diff > 1  ~ str_glue("was higher than in {comp_yr} ({.[[5]]}%)") ))

# mcv2 coverage change
cvg_mcv2 <- cvg_v %>%
  filter(vaccine == "mcv2") %>%
  mutate(diff = .[[7]] - .[[5]],
         cvg_chng_txt = case_when(
           diff %in% c(-1:1) ~ str_glue("remained relatively stable, within 1 percentage point of the {comp_yr} level"),
           diff < -1 ~ str_glue("decreased from {.[[5]]}% in {comp_yr} to {.[[7]]}% in {rev_yr}"),
           diff > 1  ~ str_glue("increased from {.[[5]]}% in {comp_yr} to {.[[7]]}% in {rev_yr}") ))

# unvaccinated mcv1 vs mcv2
unvac_mcv <- unvac_v %>%
  select(iso3c, country, vaccine, 7) %>%
  pivot_wider(names_from = vaccine, values_from = last_col()) %>%
  mutate(diff = mcv2 - mcv1) %>%
  label_vals_millions(mcv1, "mcv1_lbl") %>%
  label_vals_millions(mcv2, "mcv2_lbl") %>%
  label_vals_millions(diff, "diff_lbl")

# regional rank for mcv1
reg_rank_mcv1 <- reg_peers %>%
  filter(vaccine == "mcv1", year == rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

# dtp1->mcv1 dropout
dropout_mcv <- df %>%
  filter(year %in% c(comp_yr, rev_yr), vaccine %in% c("dtp1", "mcv1")) %>%
  select(iso3c, country, year, vaccine, vaccinated) %>%
  pivot_wider(names_from = vaccine, values_from = vaccinated) %>%
  mutate(dropout = round(((dtp1 - mcv1) / dtp1) * 100, 0),
         dropout_txt = case_when(
           dropout < 5  ~ "low DTP-MCV dropout rates imply good retention and ability to provide a full course of vaccines in infancy",
           dropout <= 10 ~ "medium DTP-MCV dropout rates imply moderate retention in immunization programmes",
           dropout > 10  ~ "high DTP-MCV dropout rates imply poor retention in immunization programmes"))

dropout_comp_mcv <- dropout_mcv %>%
  select(-c(dtp1, mcv1, dropout_txt)) %>%
  pivot_wider(names_from = year, values_from = dropout) %>%
  mutate(diff = .[[4]] - .[[3]],
         diff_txt = case_when(
           diff > 0 ~ str_glue("{diff} percentage points higher than in {comp_yr} ({.[[4]]}%)"),
           diff < 0 ~ str_glue("{abs(diff)} percentage points lower than in {comp_yr} ({.[[4]]}%)"),
           TRUE ~ str_glue("the same as in {comp_yr}")))

# persistently low mcv1 (last 3 years < 80%)
nyrs <- 3; thrshold <- 80
cvg_trend_mcv1 <- df %>%
  filter(vaccine == "mcv1", year %in% seq(rev_yr - nyrs + 1, rev_yr)) %>%
  mutate(ltx = ifelse(coverage < thrshold, 1, 0)) %>%
  summarise(sum = sum(ltx), mean_cvg = mean(coverage, na.rm = TRUE))

persistent_low_txt <- if (cvg_trend_mcv1$sum == nyrs) {
  str_glue("{ctry_name} has consistently had MCV1 coverage below {thrshold}% over the last {nyrs} years")
} else { str_glue("{ctry_name} has not consistently had MCV1 coverage below {thrshold}% over the last {nyrs} years") }

# # subnational variation for mcv1
# sub_mcv1 <- sub_df %>%
#   filter(vaccine == "mcv1", year == rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
# 
# sub_var_mcv1_txt <- case_when(
#   sub_mcv1$range_cvg >= 20 ~ str_glue("Subnational variation is high, with MCV1 coverage ranging from {sub_mcv1$min_cvg}% to {sub_mcv1$max_cvg}% across {sub_mcv1$n_admin1} admin1 areas"),
#   sub_mcv1$range_cvg >= 10 ~ str_glue("Subnational variation is moderate, with MCV1 coverage ranging from {sub_mcv1$min_cvg}% to {sub_mcv1$max_cvg}% across {sub_mcv1$n_admin1} admin1 areas"),
#   TRUE ~ str_glue("Subnational variation is low, with MCV1 coverage ranging from {sub_mcv1$min_cvg}% to {sub_mcv1$max_cvg}% across {sub_mcv1$n_admin1} admin1 areas"))
```

\textcolor{cyan}{\Large MCV: Canary in the coal mine}

\vspace{1em}

- MCV1 coverage in `r rev_yr` (`r cvg_mcv1 %>% pull(7)`%) `r cvg_mcv1 %>% pull(cvg_chng_txt)`.

\vspace{1em}

- MCV2 coverage `r cvg_mcv2 %>% pull(cvg_chng_txt)`. This leaves `r unvac_mcv$mcv1_lbl` children without any protection against measles and another `r unvac_mcv$diff_lbl` with only partial protection.

\vspace{1em}

- In `r rev_yr`, `r dropout_mcv %>% filter(year == rev_yr) %>% pull(dropout)`% of children who received DTP1 did not receive MCV1. DTP-MCV dropout was `r dropout_comp_mcv %>% pull(diff_txt)`. `r dropout_mcv %>% filter(year == rev_yr) %>% pull(dropout_txt) %>% str_to_sentence()`.

\vspace{1em}

- `r persistent_low_txt`.

\vspace{1em}

- Among `r ctry_regn` region peers, `r ctry_name` ranked `r reg_rank_mcv1$rank` out of `r n_peers` countries in MCV1 coverage in `r rev_yr`.

\vspace{1em}

\vspace{1em}

```{r setup-heatmap-mcv1-sub, include=FALSE}
# heatmap_sub_mcv1 <-
#   sub_df %>%
#   filter(vaccine == "mcv1", year >= 2000) %>%
#   mutate(year = as.factor(year)) %>%
#   ggplot(aes(x = year, y = reorder(admin1_master, coverage), fill = coverage)) +
#   geom_tile(color = "white") +
#   geom_text(aes(label = coverage), colour = "black", size = 2.8) +
#   labs(x = "", y = "",
#        title = str_glue("MCV1 coverage (%), by admin1, {ctry_name}, 2000-{rev_yr}"),
#        caption = wuenic_src) +
#   scale_x_discrete(position = "top") +
#   theme_minimal() +
#   theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
#         axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
#   scale_fill_gradientn(
#     colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
#     values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
#     guide = "colorbar", limits = c(0, 100),
#     name = "Vaccine coverage (%)  ",
#     breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

```{r ggplot-mcv1-sub, echo=FALSE, fig.width=10, fig.height=6.5}
#print(heatmap_sub_mcv1)
```

\newpage

<!-- ===================== PAGE 5: HPV ===================== -->

```{r setup-hpv, include=FALSE}
# hpv coverage change for this country
hpv <- hpv_df %>%
  select(iso3c, country, vaccine_code, year, coverage) %>%
  filter(year %in% c(hpv_rev_yr, hpv_rev_yr - 1)) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  mutate(diff = .[[5]] - .[[4]],
         chng = case_when(diff > 0 ~ "increased", diff < 0 ~ "decreased", TRUE ~ "remained constant"),
         diff_txt = case_when(
           chng %in% c("increased", "decreased") ~ str_glue("{chng} from {.[[4]]}% in {hpv_rev_yr-1} to {.[[5]]}% in {hpv_rev_yr}"),
           TRUE ~ str_glue("{chng} at {.[[5]]}% between {hpv_rev_yr-1} and {hpv_rev_yr}")),
         gt90 = case_when(
           .[[5]] < 80  & chng == "increased" ~ "Despite the increase, HPVc coverage remains below the 90% target",
           .[[5]] < 80  & chng != "increased" ~ "HPVc coverage remains below the 90% target",
           .[[5]] %in% 80:89 & chng == "increased" ~ "Despite the increase, HPVc coverage has not yet reached the 90% target",
           .[[5]] %in% 80:89 & chng != "increased" ~ "HPVc coverage has not yet reached the 90% target",
           .[[5]] == 90 ~ "Coverage of HPV last dose achieved the 90% target",
           .[[5]] > 90  ~ "Coverage of HPV last dose has exceeded the 90% target",
           TRUE ~ NA_character_))

# unvaccinated girls (hpv1 and hpvc)
hpv_unvac <- hpv_df %>%
  filter(year == hpv_rev_yr) %>%
  select(iso3c, country, vaccine_code, unvaccinated) %>%
  pivot_wider(names_from = vaccine_code, values_from = unvaccinated) %>%
  mutate(diff = PRHPVC_F - PRHPV1_F) %>%
  label_vals_millions(PRHPVC_F, "PRHPVC_F_lbl") %>%
  label_vals_millions(PRHPV1_F, "PRHPV1_F_lbl") %>%
  label_vals_millions(diff, "diff_lbl")

# whether this country has introduced hpv
hpv_intro <- wiise_hpv_intro_yrs %>%
  filter(iso3c == ctry_iso) %>%
  mutate(
    introduced = case_when(
      nationwide == "yes" | partially == "yes" ~ 1,
      TRUE ~ 0
    )
  )

hpv_intro_flag <- ifelse(nrow(hpv_intro) > 0 && any(hpv_intro$introduced == 1, na.rm = TRUE), 1, 0)

# get first year of hpv data
first_hpv_year <- hpv_intro %>% filter(introduced == 1) %>% pull(year) %>% min(na.rm = TRUE)
#last_hpv_year <- hpv_intro %>% filter(iso3c == ctry_iso) %>% pull(year) %>% max(na.rm = TRUE)

# regional rank for hpv
reg_rank_hpv <- hpv_peers %>%
  filter(vaccine_code == "PRHPVC_F", year == hpv_rev_yr) %>%
  arrange(desc(coverage)) %>%
  mutate(rank = row_number()) %>%
  filter(iso3c == ctry_iso)

n_peers_hpv <- hpv_peers %>% filter(vaccine_code == "PRHPVC_F", year == hpv_rev_yr) %>% n_distinct("iso3c")

# # subnational hpv variation (if available)
# sub_hpv <- sub_df %>%
#   filter(grepl("hpv", vaccine, ignore.case = TRUE), year == hpv_rev_yr) %>%
#   summarise(min_cvg = min(coverage, na.rm = TRUE), max_cvg = max(coverage, na.rm = TRUE),
#             range_cvg = max(coverage, na.rm = TRUE) - min(coverage, na.rm = TRUE),
#             n_admin1 = n_distinct(admin1_master))
```

\textcolor{cyan}{\Large HPV:}

\vspace{1em}

```{r hpv-intro-flag, results='asis', echo=FALSE}
# if the country has not introduced hpv, flag this and skip the rest
if (hpv_intro_flag == 0) {
  cat(str_glue("-   {ctry_name} had not introduced HPV vaccination as of {rev_yr}.\n\n"))
} else {
  cat(str_glue("-   Programme coverage of HPV1 among girls {hpv %>% filter(vaccine_code == 'PRHPV1_F') %>% pull(diff_txt)}, and coverage of the last dose (HPVc) {hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(diff_txt)}.\n\n"))
  cat("\\vspace{1em}\n\n")
  cat(str_glue("-   {hpv %>% filter(vaccine_code == 'PRHPVC_F') %>% pull(gt90)}.\n\n"))
  cat("\\vspace{1em}\n\n")
  cat(str_glue("-   In {hpv_rev_yr}, {hpv_unvac$PRHPV1_F_lbl} girls in {ctry_name} did not receive any HPV vaccination.\n\n"))
  cat("\\vspace{1em}\n\n")
  if (nrow(reg_rank_hpv) > 0) cat(str_glue("-   Among {ctry_regn} region peers, {ctry_name} ranked {reg_rank_hpv$rank} out of {n_peers_hpv} countries in HPVc coverage in {hpv_rev_yr}.\n\n"))
  # if (sub_hpv$n_admin1 > 0) {
  #   sub_var_hpv_txt <- case_when(
  #     sub_hpv$range_cvg >= 20 ~ str_glue("Subnational variation in HPV coverage is high, ranging from {sub_hpv$min_cvg}% to {sub_hpv$max_cvg}% across {sub_hpv$n_admin1} admin1 areas"),
  #     sub_hpv$range_cvg >= 10 ~ str_glue("Subnational variation in HPV coverage is moderate, ranging from {sub_hpv$min_cvg}% to {sub_hpv$max_cvg}% across {sub_hpv$n_admin1} admin1 areas"),
  #     TRUE ~ str_glue("Subnational variation in HPV coverage is low, ranging from {sub_hpv$min_cvg}% to {sub_hpv$max_cvg}% across {sub_hpv$n_admin1} admin1 areas"))
  #   cat("\\vspace{1em}\n\n")
  #   cat(str_glue("-   {sub_var_hpv_txt} in {hpv_rev_yr}.\n\n"))
  # }
}
```

\newpage

<!-- ===================== PAGE 6: REGIONAL COMPARISON ===================== -->

```{r setup-regional-comparison, include=FALSE}
# coverage for all regional peers in rev_yr, for dtp1, dtp3, mcv1, hpvc
reg_comp_vax <- reg_peers %>%
  filter(vaccine %in% c("dtp1", "dtp3", "mcv1"), year == rev_yr) %>%
  select(iso3c, country, vaccine, coverage)

reg_comp_hpv <- hpv_peers %>%
  filter(vaccine_code == "PRHPVC_F", year == hpv_rev_yr) %>%
  select(iso3c, country, coverage) %>%
  mutate(vaccine = "hpvc")

reg_comp <- bind_rows(reg_comp_vax, reg_comp_hpv) %>%
  mutate(
    highlight = ifelse(iso3c == ctry_iso, str_glue("{ctry_name}"), "Peers"),
    vaccine = toupper(vaccine),
    vaccine = factor(vaccine, levels = c("DTP1", "DTP3", "MCV1", "HPVC"))
  )

# regional average per vaccine
reg_avg <- reg_comp %>%
  group_by(vaccine) %>%
  summarise(avg = round(mean(coverage, na.rm = TRUE), 0))

plt_reg_comp <- reg_comp %>%
  mutate(country = reorder_within(country, coverage, vaccine)) %>%
  ggplot(aes(x = coverage, y = country, colour = highlight, size = highlight)) +
  geom_point() +
  geom_vline(data = reg_avg, aes(xintercept = avg), linetype = "dashed", colour = "grey50", linewidth = 0.5) +
  facet_wrap(~vaccine, ncol = 2, scales = "free_y") +
  scale_y_reordered() + 
  scale_colour_manual(values = setNames(c("#E2231A", "#A8D5E2"), c(ctry_name, "Peers")), name = "") +
  scale_size_manual(values = setNames(c(4, 2.5), c(ctry_name, "Peers")), name = "") +
  labs(x = "Coverage (%)", y = "",
       title = str_glue("Coverage comparison: {ctry_name} vs. {ctry_regn} peers, {rev_yr}"),
       subtitle = "Dashed line = regional average",
       caption = wuenic_src) +
  theme_unicef() +
  theme(legend.position = "bottom", 
        strip.text = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        axis.text.y = element_text(size = 7))
```

\textcolor{cyan}{\Large Regional Comparison}

\vspace{1em}

The chart below shows how `r ctry_name` compares to other countries in the `r ctry_regn` region on DTP1, DTP3, MCV1, and HPVc coverage in `r rev_yr`. The dashed line represents the regional average for each vaccine.

\vspace{1em}

```{r ggplot-reg-comp, echo=FALSE, fig.width=10, fig.height=7}
print(plt_reg_comp)
```

\newpage

<!-- ===================== PAGE 7: ADDITIONAL VACCINES ===================== -->

```{r setup-additional-vax, include=FALSE}
# additional vaccines to check: bcg, hib3 or hepb3, ipv1, pcv3 or rotac
# -- only include if the vaccine has non-missing coverage in this country
add_vax <- c("bcg", "hib3", "hepb3", "ipv1", "pcvc", "rotac")

add_vax_df <- df %>%
  filter(vaccine %in% add_vax, year %in% c(comp_yr, rev_yr)) %>%
  select(iso3c, country, vaccine, year, coverage) %>%
  pivot_wider(names_from = year, values_from = coverage) %>%
  filter(!is.na(.[[4]]) | !is.na(.[[5]])) %>%    # keep only vaccines with data
  mutate(diff = .[[5]] - .[[4]],
         diff_txt = case_when(
           diff > 0  ~ str_glue("increased from {.[[4]]}% in {comp_yr} to {.[[5]]}% in {rev_yr}"),
           diff < 0  ~ str_glue("decreased from {.[[4]]}% in {comp_yr} to {.[[5]]}% in {rev_yr}"),
           diff == 0 ~ str_glue("remained the same ({.[[4]]}%) between {comp_yr} and {rev_yr}"),
           TRUE ~ NA_character_),
         vaccine_label = case_when(
           vaccine == "bcg"   ~ "BCG",   vaccine == "hib3"  ~ "Hib3",
           vaccine == "hepb3" ~ "HepB3", vaccine == "ipv1"  ~ "IPV1",
           vaccine == "pcvc"  ~ "PCV3",  vaccine == "rotac" ~ "RotaC",
           TRUE ~ toupper(vaccine)),
         gt90 = ifelse(.[[5]] >= 90, "achieved", "did not achieve"))

# summary heatmap for additional vaccines over time
add_vax_heatmap_df <- df %>%
  filter(vaccine %in% add_vax, year >= 2000) %>%
  mutate(vaccine_label = case_when(
    vaccine == "bcg" ~ "BCG", vaccine == "hib3" ~ "Hib3", vaccine == "hepb3" ~ "HepB3",
    vaccine == "ipv1" ~ "IPV1", vaccine == "pcvc" ~ "PCV3", vaccine == "rotac" ~ "RotaC",
    TRUE ~ toupper(vaccine)))

plt_add_vax <-
  add_vax_heatmap_df %>%
  ggplot(aes(x = as.factor(year), y = vaccine_label, fill = coverage)) +
  geom_tile(color = "white") +
  geom_text(data = add_vax_heatmap_df %>% filter(coverage >= 60), aes(label = coverage), colour = "black", size = 3) +
  geom_text(data = add_vax_heatmap_df %>% filter(coverage < 60),  aes(label = coverage), colour = "white", size = 3) +
  labs(x = "", y = "",
       title = str_glue("Additional vaccine coverage (%), {ctry_name}, 2000-{rev_yr}"),
       caption = wuenic_src) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(limits = rev) +
  theme_minimal() +
  theme(panel.grid = element_blank(), legend.position = "bottom", legend.key.width = unit(1.2, "cm"),
        axis.text.x = element_text(angle = 45, hjust = 0, size = 9)) +
  scale_fill_gradientn(
    colours = c("#B50800", "#E2231A", "#F26A21", "#FFC20E", "#80BD41", "#00833D"),
    values  = scales::rescale(c(0, 59, 69, 79, 89, 94, 100)),
    guide = "colorbar", limits = c(0, 100),
    name = "Vaccine coverage (%)  ",
    breaks = c(0, 59, 69, 79, 89, 100), labels = c("0", "60", "70", "80", "90", "100"))
```

\textcolor{cyan}{\Large Additional Vaccines}

\vspace{1em}

```{r additional-vax-bullets, results='asis', echo=FALSE}
# dynamically generate one bullet per vaccine with data
for (i in seq_len(nrow(add_vax_df))) {
  row <- add_vax_df[i, ]
  cat(str_glue("-   {row$vaccine_label} coverage {row$diff_txt} and {row$gt90} the 90% target in {rev_yr}."), "\n\n")
  cat("\\vspace{1em}\n\n")
}
```

```{r ggplot-add-vax, echo=FALSE, fig.width=10, fig.height=4}
print(plt_add_vax)
```

\newpage

<!-- ===================== APPENDIX ===================== -->

\textcolor{cyan}{\Large Appendix}

\vspace{1em}

\textcolor{cyan}{\Large WHO/UNICEF Estimates of National Immunization Coverage (WUENIC)}

\vspace{1em}

Each year, WHO and UNICEF review national immunization data from Member States, including administrative and official coverage, survey estimates and contextual information, and use rules-based data triangulation approach to assess coverage. Estimates are produced independently, without borrowing data from other countries, and are not based on ad hoc adjustments. In some cases, they rely on a single source, typically nationally reported coverage. When data for a specific country-vaccine-year are unavailable, values from adjacent years are interpolated or extrapolated, and conflicting sources are assessed to determine the most reliable estimate, considering potential biases and expert opinion.

\vspace{0.5em}
Methods:

\vspace{0.5em}

\begin{itemize}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://pubmed.ncbi.nlm.nih.gov/19649368/}{Burton et al. 2009. WHO and UNICEF estimates of national infant immunization coverage: methods and processes}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0047806}{Burton et al. 2012. A formal representation of the WHO and UNICEF estimates of national immunization coverage: a computational logic approach}}
        \end{sloppypar}
  \item \begin{sloppypar}
        \textcolor{blue}{\href{https://openpublichealthjournal.com/VOLUME/6/PAGE/73/ABSTRACT/}{Brown et al. 2013. An introduction to the grade of confidence used to characterize uncertainty around the WHO and UNICEF estimates of national immunization coverage}}
        \end{sloppypar}
\end{itemize}

\vspace{0.5em}

These talking points present the latest WUENIC estimates (published 15 July `r rev_yr+1`).

\vspace{1.5em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont Definitions of immunization terms:}

\vspace{0.5em}

-   Bacillus Calmette-Guerin (BCG): vaccine against tuberculosis
\vspace{0.5em}
-   Hepatitis B birth dose, given within 24 hours after birth (HepBB)
\vspace{0.5em}
-   Diphtheria, tetanus, and pertussis vaccine, first dose (DTP1) and third dose (DTP3)
\vspace{0.5em}
-   Hepatitis B vaccine, third dose (HepB3)
\vspace{0.5em}
-   Haemophilus influenzae type B vaccine, third dose (Hib3)
\vspace{0.5em}
-   Poliomyelitis vaccine, third dose (Polio3)
\vspace{0.5em}
-   Inactivated polio vaccine, first dose (IPV1) and second dose (IPV2): second dose is only shown for oral polio vaccine (OPV) using countries
\vspace{0.5em}
-   Measles containing vaccine, first dose (MCV1) and second dose (MCV2)
\vspace{0.5em}
-   Rotavirus vaccine, last dose (RotaC)
\vspace{0.5em}
-   Pneumococcal vaccine, third dose (PCV3)
\vspace{0.5em}
-   Yellow Fever vaccine (YFV)
\vspace{0.5em}
-   Meningococcal A vaccine (MengA)
\vspace{0.5em}
-   Human Papillomavirus vaccine, first dose (HPV1) and last dose (HPVc)

\vspace{1.5em}

\textcolor{cyan}{\fontsize{14}{16}\selectfont Additional resources:} \textcolor{blue}{\href{https://data.unicef.org/resources/immunization/}{\underline{https://data.unicef.org/resources/immunization/}}}
